---
title: "IMDB Data Analysis"
author: Qingsong Wang
date: April 8th, 2019
output: 
  ioslides_presentation:
    widescreen: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width=6, fig.height=3) 
options(digits = 3)
```
# Where Do We Start?

## Background
- **Movies**, also known as films, are a type of visual communication which uses moving pictures and sound to tell stories or teach people something. 
- Movies are appealing a rapidly-increasing number of audience.
- The questions that I'm interested in are listed delow:
  * Which factors are of importance in making the movies popular, and how to predict the audience rating.
  * How to determine whether the investment on a movie is profiable or not.
  * How to build a recommender system for the movie audience.
  * etc
  
## Data Description
- The two datasets comes from [Kaggle](https://www.kaggle.com/), and we combine them.
- The combined dataset contains nearly $5000$ movies, and after removing NA rows $3636$ pieces are left
- The variables are listed on the next page

## Data Description
```{r, include=FALSE}
library(dplyr)
library(readr)
library(wordcloud)
library(jsonlite)
library(tidyverse)
library(ggplot2)
library(plotly)
library(lubridate)
library(gridExtra)
library(GGally)
library(network)
library(formattable)
library(DT)
library(RColorBrewer)
library(ggcorrplot)
library(randomForest)
library(e1071)
library(rpart)
movie = read_csv('movie.csv')
```
```{r}
colnames(movie)
```
## Data Description
- We explain some variables here:
  - aspect_ratio: the aspect ratio of a movie
  - actor_x_name / actor_x_facbook_likes: name and facebook like number of the actor
  - facenumber_in_poster: Contains number of actor¡¯s faces on a movie poster
  - gross: the revenue of a movie in USD
  - genres: the sub-genres to which a movie belongs
  - imdb_score: the audience rating of a movie
  - movie_title: title of a movie
  - title_year: the year in which a film was released
  
## Data Description
- Considering inflation, it's more reasonable to measure the revenue of a movie by a ratio, i.e. the return rate,
$$
\text{return_rate} = \frac{\text{gross} - \text{budget}}{\text{budget}}
$$
- We regard the **imdb_score** and **return_rate** as the response respectively and apply several machine learning techniques to predict. 
- Whether a movie is profitable or not can also be treated as a $0-1$ classification problem, so classifiction tools can be helpful.

# What Do We Want?

## Goal of Our Analysis
There are mainly three goals as listed below.

- Obtain a comprehensive knowledge of the dataset.
- Build a model to predict the IMDB score of a movie based on several variables.
- Build a model to determine whether a movie is profitable, or in other words, worth investing.

# How Do We Begin? 
```{r, include=FALSE}
f1 = list(
  family = "Old Standard TT, serif",
  size = 14,
  color = "grey"
)
f2 = list(
  family = "Old Standard TT, serif",
  size = 10,
  color = "black"
)
a = list(
  titlefont = f1,
  showticklabels = T,
  tickangle = -45,
  tickfont = f2
)

m = list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

# annotations for subplot
a1 = list(x = 0.5, y = 1.0,
           showarrow = FALSE, 
           text = "Distribution of bugdet", 
           xanchor = "center", 
           xref = "paper", 
           yanchor = "bottom", 
           yref = "paper", 
           font = f1)

b1 = list(x = 0.5, y = 1.0, 
           showarrow = FALSE, 
           text = "Distribution of gross", 
           xanchor = "center", 
           xref = "paper", 
           yanchor = "bottom", 
           yref = "paper",
           font = f1)


# creating a function called scatter_plot for
# plotting scatter plots using ggplot and plotly

scatter_plot = function(x, y, xlabel, ylabel, title,
                         text1, text2, text3,
                         alpha = NULL){
  if(is.null(alpha)) alpha = 0.4
  gp = ggplot(data = movie, mapping = aes(x = x, y = y,
                                           text = paste(text1, x,
                                                        text2, y,
                                                        text3, movie_title)))
  plot = gp + geom_point(position = "jitter", 
                          show.legend = F, shape = 21, 
                          stroke = .2, alpha = alpha) + 
    xlab(xlabel) +
    ylab(ylabel) +
    ggtitle(title) +
    theme_minimal() +
    theme(legend.position = "none",
          plot.title = element_text(size = 12, face = "bold", 
                                    family = "Times", 
                                    color = "darkgrey")) 
  
  ggplotly(plot, tooltip = "text") %>%
    layout(m, xaxis = a, yaxis = a)
}

# creating function for plotting scatter plot using facet wrap
facet_plot = function(x, y, xlabel, ylabel, alpha = NULL){
  if(is.null(alpha)) alpha = 1
  
  fp = ggplot(data = movie, mapping = aes(x = x, y = y))
  fp + geom_point(aes(fill = genres), position = "jitter", 
                  show.legend = F, shape = 21,
                  stroke = 0.2, alpha = alpha) +
    xlab(xlabel) +
    ylab(ylabel) +
    facet_wrap(~genres, scales = "free") +
    theme_minimal()
}

# creating a function for plotting a simple histogram
hist_plot = function(x, xlabel, bwidth, fill = NULL, color = NULL){
  if(is.null(fill)) fill = "orange"
  if(is.null(color)) color = "black"
  hp = ggplot(data = movie, mapping = aes(x = x))
  gp = hp + geom_histogram(binwidth = bwidth, fill = fill, 
                            color = color, 
                            size = 0.2,
                            alpha = 0.7,
                            show.legend = F) +
    xlab(xlabel) +
    theme_minimal()
  
  ggplotly(gp) %>%
    layout(margin = m, xaxis = a, yaxis = a)
}

# creating a function for plotting histogram using facet wrap
facet_hist_plot = function(x, xlabel, bwidth){
  
  hp = ggplot(data = movie, mapping = aes(x = x))
  hp + geom_histogram(aes(fill = genres), binwidth = bwidth,
                      show.legend = F, 
                      color = "black", size = 0.2, 
                      alpha = 0.8) +
    xlab(xlabel) +
    theme_minimal() +
    theme(legend.position = "none",
          axis.text = element_text(size = 12, angle = 20),
          axis.title = element_text(size = 14, 
                                    family = "Times", 
                                    color = "darkgrey",
                                    face = "bold")) + 
    facet_wrap(~ genres, scales = "free_y", ncol = 4)    
}


# creating function for plotting histograms for budget and gross
budg_gross_hist = function(x){
  
  bh = ggplot(movie, aes(x = x))
  bh + geom_histogram(binwidth = 0.05, 
                      fill = sample(brewer.pal(11, "Spectral"), 1),
                      color = "black", 
                      size = 0.09,
                      alpha = 0.7) + 
    scale_x_log10() + 
    theme_minimal()
  
  ggplotly() %>% 
    layout(m, xaxis = a, yaxis = a) 
}


# creating a function for ploting bar graphs
bar_plot = function(data, x, y, info, xlabl, ylabl, title, 
                     deci = NULL, suf = NULL){
  if(is.null(suf)) suf = ""
  if(is.null(deci)) deci = 0
  b1 = ggplot(data, aes(x = reorder(genres, x), 
                         y = y, 
                         text = paste("Genre:", genres,
                                      info, 
                                      round(y, deci), suf)))
  
  b1 + geom_bar(aes(fill = genres), stat = "identity", 
                show.legend = F, color = "black", size = 0.2,
                width = 0.7, alpha = 0.7) +
    xlab(xlabl) +
    ylab(ylabl) +
    ggtitle(title) +
    theme_minimal() +
    theme(legend.position = "none",
          plot.title = element_text(size = 14, 
                                    color = "grey", family = "Times")) +
    scale_fill_brewer(palette = "Spectral")
  
  ggplotly(tooltip = "text") %>%
    layout(margin = m, xaxis = a, yaxis = a)
}


# creating a function for plotting plotly line graph for title_year
line_graph = function(data, y, name){
  
  scat_p1 = plot_ly(data, x = ~title_year, y = ~ y, 
                     name = name, type = 'scatter', mode = 'lines',
                     line = list(color = sample(brewer.pal(11, "Spectral"), 1))) %>%
    layout(xaxis = list(title = "Title Year", zeroline = F,
                        showline = F, 
                        showticklabels = T),
           yaxis = list(title = "Average Score"),
           title = "Line Graph for Avg Score/Avg Votes/Avg User Review by Title Year", 
           font = list(family = "Serif", color = "grey"),
           legend = list(orientation = "h", size = 6,
                         bgcolor = "#E2E2E2",
                         bordercolor = "darkgrey",
                         borderwidth = 1),
           margin = m)
  scat_p1
}
colors = c(brewer.pal(n = 11, name = "Spectral"))
```

## Exploratory Data Analysis 
- Display the proportion of movies making profit or loss. 
```{r}
movie %>%
  group_by(if_profit) %>%
  summarise(count = n()) %>% 
  plot_ly(labels = ~factor(if_profit), values = ~count,
          insidetextfont = list(color = 'Black'),
          marker = list(colors = sample(brewer.pal(11, "Spectral")), 
                        line = list(color = 'Black', width = .5)), 
          opacity = 0.8) %>%
  add_pie(hole = 0.6) %>%
  layout(title = "Share of movies making profit or loss", 
         titlefont = list(family = "Times", size = 20, color = "grey"),
         xaxis = list(showgrid = T, zeroline = F, showticklabels = F),
         yaxis = list(showgrid = T, zeroline = F, showticklabels = F),
         showlegend = F, 
         margin = list(t = 50, b = 50))
```
- The number of profitable movies is a bit larger, but we can still say that the dataset is balanced.

## Exploratory Data Analysis
- Check the distribution of **imdb_score**.
```{r}
hist_plot(movie$imdb_score, bwidth = 0.1, 
          "IMDB Score", 
          fill = sample(brewer.pal(11, "Spectral"), 1),
          color = "black") 
```
- There exists right-skewness in the distribution of the IMDB score, most of which lie around $6.7$. We might take $\log$ or Box-Cox transformation to handle this problem.

## Exploratory Data Analysis
- Check the distribution of **budget** and **gross**, with the values outside their respective inter quartile ranges removed.
```{r}
options(scipen = 999)  
# transformed budget histograms
p1 = budg_gross_hist(movie$budget) %>% 
  layout(annotations = a1)

# transformed gross histograms
p2 = budg_gross_hist(movie$gross) %>%
  layout(annotations = b1) 

subplot(p1, p2, widths = c(0.5, 0.5))
```
- Similar to **imdb_score**, right-skewness exist and $\log$ transformation may still help.

## Exploratory Data Analysis
- Check if profitable movies recieve more appreciation by the users. Note that $1$ represents profitable movies and $0$ the others. 
```{r}
prof_loss = movie %>%
  ggplot(aes(x = imdb_score, color = factor(if_profit))) +
  geom_freqpoly(bins = 30) +
  scale_color_manual(values = c(sample(brewer.pal(11, "Spectral"))),
                     labels = c("Loss", "Profit")) +
  xlab("IMDB Score") +
  theme_minimal()

ggplotly(prof_loss) %>%
  layout(margin = list(t = 50, b = 100), xaxis = a, yaxis = a, 
         legend = list(orientation = "h", size = 4,
                       bgcolor = "#E2E2E2",
                       bordercolor = "darkgrey",
                       borderwidth = 1,
                       x = 0,
                       y = -0.1))
```
- Obviously, profitable movies gain higher IMDB score in general. A t-test provides more evidence but we're not going to show here.

## Exploratory Data Analysis
- Out of curiosity, we want to check if **imdb_score** is related to **release year**.
```{r}
movie$before_n_after_2000 = ifelse(movie$title_year >= 2000, 1, 0)

# plotting a histogram separated by before_n_after_2000
his = movie %>%
  ggplot(aes(x = imdb_score, fill = factor(before_n_after_2000))) +
  geom_histogram(binwidth = 0.1, 
                 color = "black",
                 position = "dodge",
                 size = 0.2,
                 alpha = 0.7) +
  xlab("IMDB Score") +
  ggtitle("Distribution of scores before and after year 2000") +
  scale_fill_manual(values = c(sample(brewer.pal(11, "Spectral")))) +
  theme_minimal() + 
  theme(plot.title = element_text(size = 14,
                                  colour = "darkgrey",
                                  family = "Times"))


ggplotly(his) %>%
  layout(margin = list(t = 50, b = 100), 
         xaxis = a, yaxis = a, 
         legend = list(orientation = "h", size = 4,
                       bgcolor = "#E2E2E2",
                       bordercolor = "darkgrey",
                       borderwidth = 1,
                       x = 0,
                       y = -0.1)) 
```
- We set year 2000 as a boundary and find that there's almost no difference in the distribution of **imdb_score** before and after that year.

## Exploratory Data Analysis
- Check the correlation between several variables
```{r}
ggpairs(movie %>% 
                   select(imdb_score, total_fb_likes, gross, num_voted_users))
```
- This step helps us pick up the most significant variables, which can be used in regression analysis.

## Exploratory Data Analysis


# Which Models Can Help?

## Split the Dataset
Notice that if we try to decide whether to invest a movie or not, some variables cannot be obtained in advance, such as **num_user_for_reviews**, **num_voted_users**, etc. Some text varaibles should be transformed into factor type, while some hardly have contribution, hence need dropping.  
```{r}
movie$genres = factor(movie$genres)
movie$color = factor(movie$color)
movie$content_rating = factor(movie$content_rating)
idx = c('actor_1_facebook_likes', 'actor_2_facebook_likes', 'actor_3_facebook_likes',
        'director_facebook_likes', 'cast_total_facebook_likes', 'aspect_ratio', 'budget',
        'color', 'facenumber_in_poster', 'imdb_score', 'runtime', 'return_rate', 'if_profit', 
        'total_fb_likes')
movie = movie[, idx]
```
As usual, we first split the dataset into training set and testing set.
```{r, echo = TRUE}
train_num = ceiling(0.8 * nrow(movie))
test_num = nrow(movie) - train_num
train_id = sample(1:nrow(movie), train_num)
train = movie[train_id, ]
test = movie[-train_id, ]
```


## Predict IMDB score
```{r}
lfit = lm(imdb_score ~ ., train[,1:12])
pred_score = predict(lfit, test)
print(paste('For IMDB score, RMSE is', 1 / test_num * sqrt(sum((pred_score - test$imdb_score)^2))))
lfit2 = lm(return_rate ~ budget + runtime + total_fb_likes + facenumber_in_poster + 
            num_critic_for_reviews + num_user_for_reviews + num_voted_users + aspect_ratio, train)
pred_rate = predict(lfit2, test)
print(paste('For return rate, RMSE is', 1 / test_num * sqrt(sum((pred_rate - test$return_rate)^2))))
```
## Linear Regression
```{r}
summary(lfit)
```
## Support Vector Machine
```{r}
svmfit = svm(imdb_score ~ budget + runtime + total_fb_likes + facenumber_in_poster + 
               num_critic_for_reviews + num_user_for_reviews + num_voted_users + title_year + aspect_ratio, train)
pred_score = predict(svmfit, test)
print(paste('For IMDB score, RMSE is', 1 / test_num * sqrt(sum((pred_score - test$imdb_score)^2))))
svmfit = svm(return_rate ~ budget + runtime + total_fb_likes + facenumber_in_poster + 
            num_critic_for_reviews + num_user_for_reviews + num_voted_users + aspect_ratio, train)
pred_rate = predict(svmfit, test)
print(paste('For return rate, RMSE is', 1 / test_num * sqrt(sum((pred_rate - test$return_rate)^2))))
```

## Random Forest
```{r}
rffit = randomForest(imdb_score ~ budget + runtime + total_fb_likes + facenumber_in_poster + 
               num_critic_for_reviews + num_user_for_reviews + num_voted_users + title_year, train)
pred_score = predict(rffit, test)
print(paste('For IMDB score, RMSE is', 1 / test_num * sqrt(sum((pred_score - test$imdb_score)^2))))
rffit = randomForest(return_rate ~ budget + runtime + total_fb_likes + facenumber_in_poster + 
            num_critic_for_reviews + num_user_for_reviews + num_voted_users, train)
pred_rate = predict(rffit, test)
print(paste('For return rate, RMSE is', 1 / test_num * sqrt(sum((pred_rate - test$return_rate)^2))))
```

## Logistic Regression
```{r}
logfit = glm(if_profit ~ budget + runtime + total_fb_likes + facenumber_in_poster + 
               num_critic_for_reviews + num_user_for_reviews + num_voted_users + title_year + aspect_ratio, train,
             family=binomial(link="logit"))
pred_if = predict(logfit, test, type = 'response') >= 0.5
print(paste('Correct prediction rate is', 1 / test_num * sum(pred_if == test$if_profit)))
```

# Summary
## Conclusions?
- Which variables are important?
- Which model to choose?
- Practical significance?
- What to do next?